Turtlebot SLAM

### 2013-01-22 trunk/Article/Article.tex
Corrected the small stuff. Did not include the reference for Scan mathcing yet.

\section{Introduction}
\label{sec:intro}
Nowadays mobile robots succeed to assist industry and society in a diverse set of problems. Transportation, search, rescue and, more specifically, automated vacuum cleaning are examples from different fields. Typically, mobile robots can move through an environment using legs or wheels. In addition they are able to gather information about their surrounding using sensors. Some even have devices similar to the human arm or hand which allows them to interact with the environment. A vacuum cleaning robot depicted in Figure \ref{fig:vacuum_cleaner} is an example of a robot which has wheels to move around. Furthermore it is equipped with a set of sensors. These sensors provide data about the location of obstacles as well as, for example, gaps in the floor the robot is moving on. To efficiently perform this task robots needs a map of its environment. For some tasks map data is available for the robot from the beginning. Provided a map the robot can locate itself using the incoming sensor data or, if available, GPS data. For tasks in unknown environments, however, the robot must rely on internal sensors and make its own map of the environment.

@@

Some approaches model the problem as a so called ``on-line" state estimation problem. The state of such a system consists of the current robot position and the map. New measurements augment and refine the estimate as they become available. Due to their incremental nature these filtering approaches are commonly referred to as on-line SLAM methods.

@@

\section{State of the art}
This section presents the state of the art of the hardware, software, and methods used for this project. Subsection \ref{subsec:ros} gives an outline about the operating system in which the product is implemented. Next, Subsection \ref{subsec:turtle} is about the TurtleBot. At last, autonomous exploration (Subsection \ref{subsec:sotaExplore}) and SLAM (Subsection \ref{subsec:sotaSlam}) 
\label{sec:sota} are defined and explained.
\subsection{ROS}
\label{subsec:ros}
The Robot Operating System (ROS) \cite{Quigley} is the environment used throughout the development of all product components for this project. As described by Quigley et al., writing software for robots is difficult since there is a large variety of possible hardware setups. This also makes code reuse non-trivial. An additional problem is the amount of code needed to acttually ``run" something on a robot. A deep stack including driver-level software, perception modules and, for example, abstract reasoning systems is required before the actual problem can be attacked. Since the necessary expertise and effort is beyond capabilities of most single researchers, an environment that features this functionality is required to achieve anything similar to the problem specified in Subsection \ref{sec:problem}.

@@

The ROS framework used in this paper is the product of trade-offs and prioritizations. The emphasis of this framework lies on large-scale integrative robotics research, making it useful in a wide variety of situations. The difference between ROS and the above mentioned frameworks, is ROS' applicability to a wide range of hardware setups.  It provides several so called stacks. Each is an assembly of code made to solve one specific problem. To run a stack or your own code on any robot it is necessary to configure the robot and the ROS framework in a way that they can communicate with each other. For most robots the configuration is straight forward thanks to the online documentation of ROS \citep{Roswiki}.  An example for a stack is the multi-robot simulation environment ``Stage". It is explained in the following paragraph.

\subsubsection{ROS' simulator: ``Stage"}
During the development process of a system running on a robot one curcial part is testing. A ROS stack called Stage enables the developers to test code on a simulated robot. Stage allows the simulated robot to move around in virtual environments. It visualizes sensor data and other robot data when required. The user can load her own maps into stage and specify the kind of noise to be applied to measurements. A second important stack needed especially for validation is RViz. It visualizes all requested data. In Figure \ref{fig:stage_and_rviz} Stage and RViz are presented next to each other. While Stage (Left) shows the real map, RViz (Right) presents the map data as seen by the robot.

The following subsection describes the robot used; the TurtleBot.
\subsection{TurtleBot}
\label{subsec:turtle}
The product of this project is designed to run on a mobile robot called TurtleBot (Figure \ref{fig:turtlebot}). It is a low-cost, customizable, personal robot kit. The base of the TurtleBot is an iRobot Create (A). It holds a battery pack and 150 degrees/second Single Axis Gyro. The model shown in Figure \ref{fig:turtlebot} uses a Microsoft Kinect sensor (B), the one used in this paper uses a laser range sensor, and a laptop (C) to run all the processes. The robot can also be customized using the mounting hardware (D). The robot runs on an open-source SDK based on ROS which integrates all the software needed to get the TurtleBot running. Some capabilities are included in ROS and can be used in the TurtleBot after minor configurations. These include mapping and navigation, for example. 
\begin{figure}[h]
	\centering
		\includegraphics[width=0.50\textwidth]{figures/turtlebot.png}
	\caption{A TurtleBot}
	\label{fig:turtlebot}
\end{figure}

One feature which is not included in ROS yet is exploration. The field is introduced in the following subsection.

\subsection{Exploration}
\label{subsec:sotaExplore}
Usually robots localize themselves and navigate through environments using a map which is provided by an external source. In this case navigating from one location to another is an easy problem. If, however, the robot does not have access to a map, navigating through an environment is problematic. The problem treated in this paper (Subsection \ref{sec:problem}) assumes an unknown environment and, hence, no map. The robot needs to be able to explore its environment without any knowledge about it. It needs to rely on the incoming laser and odometry data only. In addition to the creation of the map and the localization of itself (the SLAM problem) the robot needs to explore the environment as efficient as possible. We define a good exploration strategy as one that generates an accurate map in the shortest amount of time.

Ideally the robot updates its internal map of the environment constantly while moving. One idea to efficiently explore the world is to move to locations which maximize the robots information gain. This target location needs to be extracted from the limited information the robot has collected with its sensors.
To achieve this the robot makes use of the idea of frontier-based exploration, first introduced by Yamauchi \cite{Yamauchi} in 1998. Frontiers are defined to be the boundaries between free and unexplored space. Intuitively, when a robot moves to a frontier it is likely that it makes measurements in previously unexplored space. Given that new data can be added to its map. The explored region grows once the robot reaches a frontier. Moreover an extended map results in new frontiers. By constantly moving to frontiers the robot expands its knowledge about the environment.

The process of building an internal map representation is introduced in Subsection \ref{subsec:sotaSlam}.

@@

SLAM is a problem because, generally, localization requires a map and mapping requires the location of the device or robot that wants to create a map. There are quite a few approaches that successfully solve the SLAM problem. In general there are two interpretations of SLAM. On-line SLAM methods estimate the current robot position along with the map. The core of a system solving the on-line SLAM problem is a filtering technique. FastSlam \citep{Montemerlo02} is an example of an on-line SLAM implementation using a Particle filter to approximate the robots position and the landmark positions on the map. Other methods have the goal to find the trajectory of the robot and the map. They solve the full SLAM problem.   
An example of such a an approach is graph-based SLAM or Graph SLAM.

\subsubsection{Graph SLAM}
An intuitive way to formulate the SLAM problem is to view it as a pose graph optimization problem. Each position along the trajectory of the robot is represented by one node. Edges are labelled with relative spatial information between the connecting nodes. This graph is build incrementally. While the robot is in motion, new nodes are added to the graph. Building up the graph, as explained above, is a straight forward procedure. The essence of the Graph SLAM method, however, is the optimization step. In this step the algorithm tries to find a spatial configuration of the trajectory which is maximally consistent with a system of constraints obtained from the edges. This problem is commonly solved using least-square error minimization techniques \citep{Grisetti}.

The next section presents the algorithms implemented for this research project.

@@

In the product there exist two main packages. One deals with the autonomous exploration of the robot. The other one is responsible for map creation using Graph SLAM. Part of the autonomous exploration is performed by ROS' navigation stack move\_base. It takes care of finding a path from the current location to a specified destination. It is also capable of controlling the robot along this path. In addition to the move\_base package another external source is used by the program. The ``eigen" library \citep{eigen} provides the user with extensive linear algebra operations. Included are simple matrix operations as well as advanced techniques. Two techniques which are used in this project are Singular Value Decomposition and sparse Cholesky factorization. Both are implemented in ``eigen".

This section is structured as follows. Subsection \ref{sec:wfd} describes the implemented exploration strategy. Next, Subsection \ref{sec:implSLAM} deals with the implemented SLAM methods. This is followed by a subsection about scan matching. The section is concluded with a part about the mapping process.

\subsection{Wavefront frontier detection}
\label{sec:wfd}
One part of the assignment is autonomous exploration. The robot has to drive autonomously through an unknown world. This feature is implemented using Wavefront frontier detection (WFD) \citep{Keidar}.

WFD is a frontier-based exploration algorithm. The WFD algorithm uses four different data structures. These store map and frontier information during the search process. They are :
\begin{enumerate}
\item{Map-Open-List}
\item{Map-Close-List}
\item{Frontier-Open-List}
\item{Frontier-Close-List}
\end{enumerate}
To understand the need of these four structures it is necessary to know that WFD is using a breadth first search (BFS), starting at the current robot position. Moreover frontier detection requires the robot to have an internal representation of the current environment. In this implementation an Occupancy grid (Subsection \ref{sec:mapping}) is used.
Upon launching the algorithm it first searches for unoccupied space around the current robot location. These grid cells are saved in the Map-Open-List. 
Frontiers are found by making use of the occupancy grid's structure. The criterion for a map-point to be a frontier is the existence of both, unknown space and known unoccupied space, in adjacent map-points. When a frontier point is found a new BFS is performed. Starting at the current point it extracts all adjacent frontier points and connects them to one frontier. It is saved in the Frontier-Open-List. The benefit of using WFD is that by maintaining these lists, the algorithm only needs to incorporate points that have not been iterated yet. In fact, the algorithm keeps track of detected open space and frontiers using the above mentioned list. Following iterations make use of this information because they do not need to consider previously known regions any more.

Exploration does not only consist of finding frontiers. Based on the frontiers a target location for the robot needs to be selected. Navigating to frontier points directly is the simplest frontier-based exploration strategy. For this project, however, a more sophisticated implementation is chosen. Instead of sending the robot directly to a frontier point, a small search is performed to find suitable destinations. Generally, a small region around the frontier cell is searched. The goal is to find a cell with as much free space around it as possible. For this purposes an evaluation function has been implemented. It incorporates three features that are used to send a goal to the robot's navigation stack. The three features are :
\begin{enumerate}
\item{Distance between frontier and robot location}
\item{Number of obstacles around the designated point} \item{Distance to the previous goal}
\end{enumerate}
The first two features have equal weights, whereas the third gets a weight of 5. This configuration turned out to create the best results (see Section \ref{sec:exp}) Their weighted sum is used to find the final goal to be sent to the navigation stack. This evaluation function enables the algorithm to select a reachable destination for the robot in almost every case. Additionally the information gain is still next to optimal since the selected goal is preferred to be close to frontier points. At last the third feature makes sure that the robots path is smooth. Selecting a goal near the previous one makes the robot avoid to many loops in the emerging graph.
\subsection{GraphSLAM}
\label{sec:implSLAM}
Two different approaches of graph-based SLAM are implemented for this project. They differ in complexity and thereby applicability to the problem. Both algorithms create an occupancy grid map. 
\begin{figure}[h]
	\centering
		\includegraphics[width=0.50\textwidth]{figures/graph.png}
	\caption{An example of a simple graph (based on \citep{Grisetti}).}
	\label{fig:turtlebotexample_graph}
\end{figure}
First, an algorithm as suggested in \citep{Grisetti} is implemented. Here the graph consists of nodes representing positions along the robot trajectory. Odometry and sensor measurements are incorporated into the constraints contained in the edges. Similar to the graphSLAM procedure explained in Section \ref{subsec:sotaSlam}, an edge connects two adjacent poses of the robot. This implementation optimizes the pose graph according to the constraints. This ``optimized" trajectory then determines the positions of obstacles on the map by refining the laser scan data according to the poses they were obtained at. Each edge thus is labelled with the spatial difference between two poses \(x_i\) and \(x_j\). Let this error be denoted by \(e_{ij}(x_i, x_j)\). This information is obtained from a combination of odometry data and laser range sensor data. In fact, the odometry is corrected using a laser scan matcher (see Subsection \ref{sec:scan}).

This graph-based SLAM approach calculates the most likely path as follows:
First it finds the Jacobians \(A\) and \(B\) of every two estimated positions connected by an edge.
\begin{equation}
\label{eq:jacobi_a}
	A_{ij} = \frac{\partial e_{ij}(x)}{\partial x_i}\uplus
\end{equation}
\begin{equation}
\label{eq:jacobi_b}
	B_{ij} = \frac{\partial e_{ij}(x)}{\partial x_j}
\end{equation}
Note that in Equations \ref{eq:jacobi_a} and \ref{eq:jacobi_b} the x represents the current guess of the concerning robot position. As explained above, \(e_{ij}(x_i, x_j)\) is the error between the two poses \(x_i\) and \(x_j\). The Jacobians \(A\) and \(B\) are then added to an information matrix \(H\) (Equation \ref{eq:H}).
\begin{equation}
\label{eq:H}
\begin{aligned}
	H_{[ii]} += A^{T}_{ij} \Omega_{ij} A_{ij} \\
	H_{[ij]} += A^{T}_{ij} \Omega_{ij} B_{ij} \\
	H_{[ji]} += B^{T}_{ij} \Omega_{ij} A_{ij} \\
	H_{[jj]} += B^{T}_{ij} \Omega_{ij} B_{ij} \\
\end{aligned}
\end{equation}
In this equation $\Omega_{ij}$ is the information matrix of a measurement between \(x_i\) and \(x_j\).
\\
Furthermore a coefficient vector \(b\) is always updated as shown in Equation \ref{eq:b}. This vector in combination with the \(H\) matrix form a set of linear constraints.
\begin{equation}
\label{eq:b}
\begin{aligned}
	b_{[i]} += A^{T}_{ij} \Omega_{ij} e_{ij} \\
	b_{[j]} += B^{T}_{ij} \Omega_{ij} e_{ij} 
\end{aligned}
\end{equation}
The optimization of the vector \(x = (x_1, ..., x_n)^T \) according to the constraints specified in \(H\) and \(b\) is the core of this approach. This step is performed using sparse Cholesky factorization.

A new node is added to the graph once the robot moves 0.5 meters. The algorithm optimizes the trajectory like explained once a new node is added.\\
In one special case the algorithm adds an extra edge, connecting to non-adjacent nodes, to the graph. This happens when the distance between the two connected robot positions is smaller than a margin. In fact, this special case enables the algorithm to create loops in the graph and more accurately estimate the robots trajectory and the map. This procedure is referred to as ``loop closing". 
\\
\\
The second implementation is based on \cite{Thrun}. It creates nodes for each odometry measurement of the robot as well as for each feature in the map. Each occupancy grid cell containing an obstacle is considered a map feature. This is because all information for the occupancy-grid map is inherited in these features. Edges are labeled with spatial constraints of the two nodes they are connecting. The output of this implementation is a vector containing the positions of the landmarks as well as the trajectory of the robot. Here the constraints labeling the edges are created using Jacobians similar to the previous approach. The major difference lies in the node creation. Here nodes are added for landmarks while the previous implementation considered the robot's trajectory only. An additional difference lies in the fact that this implementation uses features to characterize the map. The algorithm needs to align each new observed feature with the set of features already seen. This matching step has to be performed and repeated along with the complete optimization procedure until convergence.

\subsection{Scan matching}
\label{sec:scan}
The scan matcher, used by this program, is necessary to acquire an optimized movement vector between two robot positions. This is needed by the Graph SLAM algorithm (see Subsection \ref{sec:implSLAM}). The scan-matcher realized for this project can be divided into two steps. First, the two scans are matched making use of raw odometry data, retrieved by the robot's motion sensor. This is improved upon by running an iteration of the iterative closest point algorithm (ICP). It tries to align the two scans minimizing the total distance of each point in the first scan to its nearest neighbor in the second.

The general procedure of the algorithm is to first find the nearest neighboring point in the second scan for all point in the first. To achieve this nearest neighbor search is used.
In this implementation multiple points can have the same corresponding neighbor-point. If there is no point from the other scan within a specified radius of one point, it is excluded. These outliers would influence further calculations negatively.
The next step in the process is to calculate the centroids of both clouds. These are subtracted from every point in the concerning cloud. In that way the clouds are centered at the origin. The calculated points are then added to two separate $2 \times n$ matrices, where $n$ is the number of neighbour pairs. Next the rotation matrix which transforms one cloud to the other is needed. The Singular Value Decomposition (SVD) of the product of the two previously mentioned matrices returns three new matrices $U$, $\Sigma$ and $V$. The required rotation matrix R can be obtained as described in Equation \ref{eq:RVUT}.
\begin{equation}
\label{eq:RVUT}
 R = V * U^T
\end{equation}
The translation vector $t$ is the difference between the two centroids. $R$ and $t$ together form a rigid transformation. This transformation matches the two scans. Moreover it is used to correct the robot's odometry data.

@@

The map is internally represented as an occupancy grid. It consists of cells which hold information about obstacles, free/- and unknown space. Using ROS' visualization program, RViz, the map can be visualized on the screen.

The occupancy grid is a structure that is represented by a one-dimensional array. The size of an occupancy grid has to be tuned problem specific. A too large grid can lead to performance problems. If the grid is too small, however, the map will be inaccurate. Fitting a map on a smaller grid results in a low resolution.
Internally the program represents the grid as a two dimensional array. This makes navigating to either side more intuitive. Each grid cell contains an integer value. Different values represent the different states a cell can be in. The value-range for ROS' occupancy grid lies between $-1$ and $100$ \cite{occupancy}. Each value has a specific meaning:
\begin{description}
\item{$-1$} : Unknown space, not yet covered by robot's perception sensors
\item{$0$} : Known free space, robot can drive on it safely
\item{$1-100$} : Probability of being an obstacle, $1$ being a low probability and $100$ being a definite obstacle.
\end{description}
\begin{figure}[htbp]
	\centering
		\includegraphics[width=0.45\textwidth]{figures/Occup.png}
	\caption{A visualized occupancy grid.}
	\label{fig:Occupancy}
\end{figure}
Figure \ref{fig:Occupancy} shows an example occupancy grid. The robot is standing in the middle. It senses the state of all cells within a distance of one. The border cells are marked with a $-1$, which means that they are still unexplored. The remaining cells are explored being either an obstacle or free space.
The occupancy grid is created using laser scan data. Each laser scan contains information about whether or not an obstacle has been detected at a certain bearing. Maximum range readings occur when the robot senses only free space in a certain direction. All laser scans are transformed into point clouds. The absolute locations of the points are dependent on the position at which the laser scan was recorded.
In the occupancy grid detected obstacle points are indicated by value 100.
Free space is being incorporated into the map by filling regions between detected points and the concerning robot position. This is achieved with the help of a linear function.

>>>>

\section{Introduction}
\label{sec:intro}
Nowadays mobile robots succeed to assist industry and society in a diverse set of problems. Transportation, search, rescue and, more specifically, automated vacuum cleaning are examples from different fields. Typically, mobile robots can move through an environment using legs or wheels. In addition they are able to gather information about their surrounding using sensors. Some even have devices similar to the human arm or hand which allows them to interact with the environment. The vacuum cleaning robot depicted in Figure \ref{fig:vacuum_cleaner} is an example of a robot which has wheels to move around. Furthermore it is equipped with a set of sensors. These sensors provide data about the location of obstacles as well as, for example, gaps in the floor the robot is moving on. To efficiently perform this task robots needs a map of its environment. For some tasks map data is available for the robot from the beginning. Provided a map the robot can locate itself using the incoming sensor data or, if available, GPS data. For tasks in unknown environments, however, the robot must make its own map of the environment.

@@

Some approaches model the problem as a so called ``on-line" state estimation problem. The state of such a system consists of the current robot pose and the map. New measurements augment and refine the estimate as they become available. Due to their incremental nature these filtering approaches are commonly referred to as on-line SLAM methods.

@@

\section{State of the art}
This section presents the state of the art of the hardware, software, and methods used for this project. Subsection \ref{subsec:ros} gives an outline about the operating system in which the product is implemented. Next, Subsection \ref{subsec:turtle} is about the TurtleBot. At last, autonomous exploration (Subsection \ref{subsec:sotaExplore}) and SLAM (Subsection \ref{subsec:sotaSlam}) 
\label{sec:sota} are defined and explained.
\subsection{ROS}
\label{subsec:ros}
The Robot Operating System (ROS) \cite{Quigley} is the environment used throughout the development of all product components for this project. As described by Quigley et al., writing software for robots is difficult since there is a large variety of possible hardware set-ups. This also makes code reuse non-trivial. An additional problem is the amount of code needed to actually ``run" something on a robot. A deep stack including driver-level software, perception modules and, for example, abstract reasoning systems is required before the actual problem can be attacked. Since the necessary expertise and effort is beyond capabilities of most single researchers, an environment that features this functionality is required to achieve anything similar to the problem specified in Subsection \ref{sec:problem}.

@@

The ROS framework used in this paper is the product of trade-offs and prioritizations. The emphasis of this framework lies on large-scale integrative robotics research, making it useful in a wide variety of situations. The difference between ROS and the above mentioned frameworks, is ROS' applicability to a wide range of hardware set-ups. It provides several so called stacks. Each is an assembly of code made to solve one specific problem. To run a stack or your own code on any robot it is necessary to configure the robot and the ROS framework in a way that they can communicate with each other. For most robots the configuration is straight forward thanks to the on-line documentation of ROS \citep{Roswiki}.  An example for a stack is the multi-robot simulation environment ``Stage". It is explained in the following paragraph.

\subsubsection{ROS' simulator: ``Stage"}
During the development process of a system running on a robot one crucial part is testing. A ROS stack called Stage enables the developers to test code on a simulated robot. Stage allows the simulated robot to move around in virtual environments. It visualizes sensor data and other robot data when required. The user can load her own maps into stage and specify the kind of noise to be applied to measurements. A second important stack needed especially for validation is RViz. It visualizes all requested data. In Figure \ref{fig:stage_and_rviz} Stage and RViz are presented next to each other. While Stage (Left) shows the real map, RViz (Right) presents the map data as seen by the robot.

\subsection{TurtleBot}
\label{subsec:turtle}
The product of this project is designed to run on a mobile robot called TurtleBot (Figure \ref{fig:turtlebot}). It is a low-cost, customizable, personal robot kit. The base of the TurtleBot is an iRobot Create (A). It holds a battery pack and a 150 degrees/second Single Axis Gyro. The model shown in Figure \ref{fig:turtlebot} uses a Microsoft Kinect sensor (B), the one used in this paper uses a laser range sensor and a laptop (C) to run all the processes. The robot can also be customized using the mounting hardware (D). The robot runs on an open-source SDK based on ROS which integrates all the software needed to get the TurtleBot running. Some capabilities are included in ROS and can be used in the TurtleBot after minor configurations. These include mapping and navigation, for example. 
\begin{figure}[h]
	\centering
		\includegraphics[width=0.50\textwidth]{figures/turtlebot.png}
	\caption{A TurtleBot}
	\label{fig:turtlebot}
\end{figure}

\subsection{Exploration}
\label{subsec:sotaExplore}
Usually robots localize themselves and navigate through environments using a map which is provided by an external source. In this case navigating from one location to another is an easy problem. If the robot does not have access to a map, navigating through an environment is problematic. The problem treated in this paper (Subsection \ref{sec:problem}) assumes an unknown environment and, hence, no map. The robot needs to be able to explore its environment without any knowledge about it. It needs to rely on the incoming laser and odometry data only. In addition to the creation of the map and the localization of itself (the SLAM problem) the robot needs to explore the environment as efficient as possible. We define a good exploration strategy as one that generates an accurate map in the shortest amount of time.

Ideally the robot updates its internal map of the environment constantly while moving. One idea to efficiently explore the world is to move to locations which maximize the robot's information gain. This target location needs to be extracted from the limited information the robot has collected with its sensors.
To achieve this the robot makes use of the idea of frontier-based exploration, first introduced by Yamauchi \cite{Yamauchi} in 1998. Frontiers are defined to be the boundaries between free and unexplored space. Intuitively, when a robot moves to a frontier it is likely that it makes measurements in previously unexplored space. Given that new data can be added to its map. The explored region grows once the robot reaches a frontier. Moreover an extended map results in new frontiers. By constantly moving to frontiers the robot expands its knowledge about the environment.

@@

SLAM is a problem because localization requires a map and mapping requires the location of the device or robot that wants to create a map. There are quite a few approaches that successfully solve the SLAM problem. In general there are two interpretations of SLAM. On-line SLAM methods estimate the current robot pose along with the map. The core of a system solving the on-line SLAM problem is a filtering technique. FastSlam \citep{Montemerlo02} is an example of an on-line SLAM implementation using a Particle Filter to approximate the robots pose and the landmark positions on the map. Other methods have the goal to find the trajectory of the robot and the map. They solve the full SLAM problem.   
An example of such a an approach is graph-based SLAM or Graph SLAM.

\subsubsection{Graph SLAM}
An intuitive way to formulate the SLAM problem is to view it as a pose graph optimization problem. Each pose along the trajectory of the robot is represented by one node. Edges are labeled with relative spatial information between the connecting nodes. This graph is built incrementally. While the robot is in motion, new nodes are added to the graph. Building up the graph, as explained above, is a straight forward procedure. The essence of the Graph SLAM method, however, is the optimization step. In this step the algorithm tries to find a spatial configuration of the trajectory which is maximally consistent with a system of constraints obtained from the edges. This problem is commonly solved using least-square error minimization techniques \citep{Grisetti}.

@@

%This is the new part
In the product there exist two main packages. One deals with the autonomous exploration of the robot. The other one is responsible for map creation using Graph SLAM. Part of the autonomous exploration is performed by ROS' navigation stack move\_base. It takes care of finding a path from the current location to a specified destination. It is also capable of controlling the robot along this path. In addition to the move\_base package another external source is used by the program. The ``Eigen" library \citep{eigen} provides the user with extensive linear algebra operations. Included are simple matrix operations as well as advanced techniques. Two techniques which are used in this project are Singular Value Decomposition and sparse Cholesky factorization. Both are implemented in ``Eigen".

This section is structured as follows. Subsection \ref{sec:wfd} describes the implemented exploration strategy. Next, Subsection \ref{sec:implSLAM} deals with the SLAM method. This is followed by a subsection about scan matching. The section is concluded with a part about the mapping process. 
%End of the new part

\subsection{Wavefront frontier detection}
\label{sec:wfd}
One part of the assignment is autonomous exploration. The robot has to drive autonomously through an unknown world. This feature is implemented using Wavefront frontier detection (WFD).

WFD is a frontier-based exploration algorithm \citep{Keidar}. It uses four different data structures. These store map and frontier information during the search process. They are:
\begin{enumerate}
\item{Map-Open-List}
\item{Map-Close-List}
\item{Frontier-Open-List}
\item{Frontier-Close-List}
\end{enumerate}
To understand the need of these four structures it is necessary to know that WFD is using a breadth first search (BFS), starting at the current robot position. Moreover frontier detection requires the robot to have an internal representation of the current environment. In this implementation an occupancy grid (Subsection \ref{sec:mapping}) is used.
Upon launching the algorithm it first searches for unoccupied space around the current robot location. These grid cells are saved in the Map-Open-List. 
Frontiers are found by making use of the occupancy grid's structure. The criterion for a map-point to be a frontier is the existence of both, unknown space and known unoccupied space, in adjacent map-points. When a frontier point is found a new BFS is performed. Starting at the current point it extracts all adjacent frontier points and connects them to one frontier. It is saved in the Frontier-Open-List. The benefit of using WFD is that by maintaining these lists, the algorithm only needs to incorporate points that have not been iterated yet. In fact, the algorithm keeps track of detected open space and frontiers using the above mentioned list. Following iterations make use of this information because they do not need to consider previously known regions any more.

Exploration does not only consist of finding frontiers. Based on the frontiers a target location for the robot needs to be selected. Navigating to frontier points directly is the simplest frontier-based exploration strategy. For this project, however, a more sophisticated implementation is chosen. Instead of sending the robot directly to a frontier point, a small search is performed to find suitable destinations. Generally, a small region around the frontier cell is searched. The goal is to find a cell with as much free space around it as possible. For this purposes an evaluation function has been implemented. It incorporates three features that are used to send a goal to the robot's navigation stack. The three features are :
\begin{enumerate}
\item{Number of obstacles in the area}
\item{Distance to the previous goal}
\item{Distance between frontier and robot location}
\end{enumerate}
The first two features have equal weights, whereas the third gets a weight of 5. This configuration turned out to create the best results (see Subsection \ref{sec:navEval}) Their weighted sum is used to find the final goal to be sent to the navigation stack. This evaluation function enables the algorithm to select a reachable destination for the robot in almost every case. Additionally the information gain is still next to optimal since the selected goal is preferred to be close to frontier points. At last the third feature makes sure that the robots path is smooth. Selecting a goal near the previous one makes the robot avoid to many loops in the emerging graph.
\subsection{Graph SLAM}
\label{sec:implSLAM}
Two different approaches of graph-based SLAM are implemented for this project. They differ in complexity and thereby applicability to the problem. Both algorithms create an occupancy grid map. 
\begin{figure}[h]
	\centering
		\includegraphics[width=0.50\textwidth]{figures/graph.png}
	\caption{An example of a simple graph (based on \citep{Grisetti}).}
	\label{fig:turtlebotexample_graph}
\end{figure}
First, an algorithm as suggested in \citep{Grisetti} is implemented. Here the graph consists of nodes representing poses along the robot trajectory. Odometry and sensor measurements are incorporated into the constraints contained in the edges. Similar to the Graph SLAM procedure explained in Section \ref{subsec:sotaSlam}, an edge connects two adjacent poses of the robot. This implementation optimizes the pose graph according to the constraints. This ``optimized" trajectory then determines the positions of obstacles on the map by refining the laser scan data according to the poses they were obtained at. Each edge thus is labeled with the spatial difference between two poses \(x_i\) and \(x_j\). Let this error be denoted by \(e_{ij}(x_i, x_j)\). This information is obtained from a combination of odometry data and laser range sensor data. In fact, the odometry is corrected using a laser scan matcher (see Subsection \ref{sec:scan}).

This graph-based SLAM approach calculates the most likely path as follows:
First it finds the Jacobians \(A\) and \(B\) of every two estimated poses connected by an edge.
\begin{equation}
\label{eq:jacobi_a}
	A_{ij} = \frac{\partial e_{ij}(x)}{\partial x_i}
\end{equation}
\begin{equation}
\label{eq:jacobi_b}
	B_{ij} = \frac{\partial e_{ij}(x)}{\partial x_j}
\end{equation}
Note that in Equations \ref{eq:jacobi_a} and \ref{eq:jacobi_b} the x represents the current guess of the concerning robot pose. As explained above, \(e_{ij}(x_i, x_j)\) is the error between the two poses \(x_i\) and \(x_j\). The Jacobians \(A\) and \(B\) are then added to an information matrix \(H\) (Equation \ref{eq:H}).
\begin{equation}
\label{eq:H}
\begin{aligned}
	H_{[ii]} += A^{T}_{ij} \Omega_{ij} A_{ij} \\
	H_{[ij]} += A^{T}_{ij} \Omega_{ij} B_{ij} \\
	H_{[ji]} += B^{T}_{ij} \Omega_{ij} A_{ij} \\
	H_{[jj]} += B^{T}_{ij} \Omega_{ij} B_{ij} \\
\end{aligned}
\end{equation}
In this equation $\Omega_{ij}$ is the information matrix of a measurement between \(x_i\) and \(x_j\).
\\
Furthermore a coefficient vector \(b\) is always updated as shown in Equation \ref{eq:b}. This vector in combination with the \(H\) matrix form a set of linear constraints.
\begin{equation}
\label{eq:b}
\begin{aligned}
	b_{[i]} += A^{T}_{ij} \Omega_{ij} e_{ij} \\
	b_{[j]} += B^{T}_{ij} \Omega_{ij} e_{ij} 
\end{aligned}
\end{equation}
The optimization of the vector \(x = (x_1, ..., x_n)^T \) according to the constraints specified in \(H\) and \(b\) is the core of this approach. This step is performed using sparse Cholesky factorization.

A new node is added to the graph once the robot moves 0.5 meters. The algorithm optimizes the trajectory like explained once a new node is added.\\
In one special case the algorithm adds an extra edge, connecting to non-adjacent nodes, to the graph. This happens when the distance between the two connected robot poses is smaller than a margin. In fact, this special case enables the algorithm to create loops in the graph and more accurately estimate the robots trajectory and the map. This procedure is referred to as ``loop closing" (see Subsection \ref{sec:loopClosing}). 
\\
\\
The second implementation is based on \cite{Thrun}. It creates nodes for each odometry measurement of the robot as well as for each feature in the map. Each occupancy grid cell containing an obstacle is considered a map feature. This is because all information for the occupancy-grid map is inherited in these features. Edges are labeled with spatial constraints of the two nodes they are connecting. The output of this implementation is a vector containing the positions of the landmarks as well as the trajectory of the robot. Here the constraints labeling the edges are created using Jacobians similar to the previous approach. The major difference lies in the node creation. Here nodes are added for landmarks while the previous implementation considered the robot's trajectory only. An additional difference lies in the fact that this implementation uses features to characterize the map. The algorithm needs to align each new observed feature with the set of features already seen. This matching step has to be performed and repeated along with the complete optimization procedure until convergence.

\subsection{Scan matching}
\label{sec:scan}
The scan matcher, used by this program, is necessary to acquire an optimized movement vector between two robot poses. This is needed by the Graph SLAM algorithm (see Subsection \ref{sec:implSLAM}). The scan-matcher realized for this project can be divided into two steps. First, the two scans are matched making use of raw odometry data, retrieved by the robot's motion sensor. This is improved upon by running an iteration of the iterative closest point algorithm (ICP). It tries to align the two scans minimizing the total distance of each point in the first scan to its nearest neighbor in the second.

The general procedure of the algorithm is to first find the nearest neighboring point in the second scan for all point in the first. To achieve this nearest neighbor search is used.
In this implementation multiple points can have the same corresponding neighbor-point. If there is no point from the other scan within a specified radius of one point, it is excluded. These outliers would influence further calculations negatively.
The next step in the process is to calculate the centroids of both clouds. These are subtracted from every point in the concerning cloud. In that way the clouds are centered at the origin. The calculated points are then added to two separate $2 \times n$ matrices, where $n$ is the number of neighbor pairs. Next the rotation matrix which transforms one cloud to the other is needed. The Singular Value Decomposition (SVD) of the product of the two previously mentioned matrices returns three new matrices $U$, $\Sigma$ and $V$. The required rotation matrix R can be obtained as described in Equation \ref{eq:RVUT}.
\begin{equation}
\label{eq:RVUT}
 R = V * U^T
\end{equation}
The translation vector $t$ is the difference between the two centroids. $R$ and $t$ together form a rigid transformation. This transformation matches the two scans. Moreover it is used to correct the robot's odometry data.

@@

The map is internally represented as an occupancy grid. It consists of cells which hold information about obstacles, free/- and unknown space.

The occupancy grid is a structure that is represented by a one-dimensional array. The size of an occupancy grid has to be tuned problem specific. A too large grid can lead to performance problems. If the grid is too small, however, the map will be inaccurate. Fitting a map on a smaller grid results in a low resolution.
Internally the program represents the grid as a two dimensional array. This makes navigating to either side more intuitive. Each grid cell contains an integer value. The value-range for ROS' occupancy grid lies between $-1$ and $100$ \cite{occupancy}. The values indicate the state as follows:
\begin{description}
\item{$-1$} : Unknown space, not yet covered by robot's perception sensors
\item{$0$} : Known free space, robot can drive on it safely
\item{$1-100$} : Probability of being an obstacle, $1$ being a low probability and $100$ being a definite obstacle.
\end{description}
\begin{figure}[htbp]
	\centering
		\includegraphics[width=0.45\textwidth]{figures/Occup.png}
	\caption{A visualized occupancy grid}
	\label{fig:Occupancy}
\end{figure}
Figure \ref{fig:Occupancy} shows an example occupancy grid. The robot stands in the center of the it. It senses the state of all cells within a distance of one. The border cells are marked with a $-1$, which means that they are still unexplored. The remaining cells are explored being either an obstacle (100) or free space (0).
The occupancy grid is created using laser scan data. Each laser scan contains information about whether or not an obstacle has been detected at a certain bearing. Maximum range readings occur when the robot senses only free space in a certain direction. All laser scans are transformed into point clouds. The absolute locations of the points are dependent on the pose at which the laser scan was recorded.
In the occupancy grid detected obstacle points are indicated by a value of 100.
Free space is being incorporated into the map by filling regions between detected points and the concerning robot position.
